---
title: "Linear Regression Analysis: Can LOS be explained by admission stage, location, and demographics?"
author: "Steph Reynolds (Stephanie.Reynolds@ucsf.edu)"
date: "`r format(Sys.time(), '%b %d, %Y  %H:%M %p')`"
output:
  html_document:
    toc: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Background

## Project

MINDSCAPE: Modeling of infectious network dynamics for surveillance, control and prevention enhancement

## Description

This script generates multivariate linear regression models to evaluate the extent to which length of stay can be explained by admission stage, location/facility, and demographics (including age, sex, race, and ethnicity). Model parameters include:

-   `LOS` -\> Length of stay that patient was hospitalized
    -   Discharge date - admission date (in days)
    -   LOS is not normally distributed (very long right tail), so most likely needs to be transformed (consider log, square root, cubic root, or rank scale)
-   `stage_adm` -\> Admission stage
    -   Also should consider categorizing admission stage into moderate (stages 4-5) and severe (stages 6+)
-   `age` -\> Age (in years) of patient at admission
-   `sex` -\> Biological sex of patient (male, female)
-   `race` -\> Race categorized into: Black, White, Asian, Hawaiian/Pacific Islander, Other, Unknown/Declined)
    -   Still unsure how to treat `race`
-   `ethnicity` -\> Ethnicity categorized into: Hispanic/Latino, Not Hispanic, Unknown/Declined

Each model is checked to meet the assumptions of linear regression and evaluated by its performance relative to other models.

## Source Data

-   UTSW: `ut_df_unique_pts.rds`
-   UCSF: `dm_stg_unique_pt_11.08.21.rds`

# Load required packages

```{r Load required packages, include=F}
library(tidyverse)
library(tidymodels)
library(performance) # check model assumptions and performance
library(vip) # calculate and visualize feature importance
library(polspline) # polymars 
library(jtools) # summarize and plot regression model results 
```


# Import and preview data

```{r Import and preview data}
regdf <- readRDS("/Users/sreynolds2/Documents/GitHub/MS-UTSW_Covid_Staging/data/regdf_11.08.21.rds")

# Check structure and preview data 
str(regdf)

# Import and preview data; keep only relevant variables
glimpse(regdf)
```

# Fit LOS to linear regression model including all variables -- FULL MODEL

LOS \~ stage_adm + loc + age + race + ethnicity + sex

```{r Linear regression model m1}
# Fit model
m1 <- lm(t.LOS ~ stage_adm + loc + age + race + ethnicity + sex, 
         data = regdf)
# Should we treat age as polynomial? -> poly(age, 2)

# Summarize regression output
summary(m1)

# Check assumptions
check_model(m1)

# Check performance
model_performance(m1)
```

# Fit LOS to linear regression model including all variables except for stage_adm

LOS \~ loc + age + race + ethnicity + sex

```{r Linear regression model m2}
# Fit model
m2 <- lm(t.LOS ~ loc + age + race + ethnicity + sex, 
         data = regdf)

# Summarize regression output
summary(m2)

# Check assumptions 
check_model(m2)

# Check performance
model_performance(m2)
```

# Fit LOS to linear regression model including all variables except for loc

LOS \~ stage_adm + age + race + ethnicity + sex

```{r Linear regression model m3}
# Fit model
m3 <- lm(t.LOS ~ stage_adm + age + race + ethnicity + sex, 
         data = regdf)

# Summarize regression output
summary(m3)

# Check assumptions
check_model(m3)

# Check performance
model_performance(m3)
```

# Fit LOS to linear regression model including only stage_adm, loc, and age

LOS \~ stage_adm + loc + age

```{r Linear regression model m4}
# Fit model
m4 <- lm(t.LOS ~ stage_adm + loc + age, 
         data = regdf)
# Should we treat age as polynomial? -> poly(age, 2)

# Summarize regression output
summary(m4)

# Check assumptions
check_model(m4)

# Check performance
model_performance(m4)

# jtools package 
summ(m4, robust = T, confint = T, vifs = T)

plot_summs(m4, scale = TRUE)

```

# Compare model coefficients for m1-m4 visually

```{r Compare model coefficients visually}
plot_summs(m1, m2, m3, m4, scale = TRUE)
```

# Compare regression output for m1-m4

```{r Compare regression output for m1-m4, results = "asis"}
export_summs(m1, m2, m3, m4, 
             scale = TRUE,
             error_format = "95% CI [{conf.low}, {conf.high}]")
```

# Univariate regression models

Predictors include: stage_adm + loc + age + race + ethnicity + sex

```{r Univariate regression models}
# ADMISSION STAGE 
m.admstage <- lm(t.LOS ~ stage_adm, data = regdf)

  # Summarize regression output
  summary(m.admstage)
  
  # Check assumptions
  check_model(m.admstage)
  
  # Check performance
  model_performance(m.admstage)
  

# LOCATION/FACILITY 
m.loc <- lm(t.LOS ~ loc, data = regdf)

  # Summarize regression output
  summary(m.loc)
  
  # Check assumptions
  check_model(m.loc)
  
  # Check performance
  model_performance(m.loc)
 

# AGE
m.age <- lm(t.LOS ~ age, data = regdf)

  # Summarize regression output
  summary(m.age)
  
  # Check assumptions
  check_model(m.age)
  
  # Check performance
  model_performance(m.age) 
  

# RACE
m.race <- lm(t.LOS ~ race, data = regdf)

  # Summarize regression output
  summary(m.race)
  
  # Check assumptions
  check_model(m.race)
  
  # Check performance
  model_performance(m.race)  
  

# ETHNICITY
m.ethnicity <- lm(t.LOS ~ ethnicity, data = regdf)

  # Summarize regression output
  summary(m.ethnicity)
  
  # Check assumptions
  check_model(m.ethnicity)
  
  # Check performance
  model_performance(m.ethnicity)  
  
  
# SEX
m.sex <- lm(t.LOS ~ sex, data = regdf)

  # Summarize regression output
  summary(m.sex)
  
  # Check assumptions
  check_model(m.sex)
  
  # Check performance
  model_performance(m.sex)

```

# Results from Univariate Regression Analyses

-   Which variables had a p-value of \< 0.2?
    -   Admission stage
    -   Location/facility
    -   Age
    -   Sex

# Important Notes

-   Compare R\^2 for all models. Which has the highest?
-   What about comparing AIC?
-   How to deal with age? Treat it as polynomial?
-   Other methods to consider:
    -   Transform LOS on log, square root, cubic root, or rank scale
    -   Rank regression
    -   Robust regression

# Other Notes from 1/21 Call with Travis (with first attempt at regression)

-   Decided a-priori that we would include stage of admission, age, and race.
-   Used fastDummies::dummy_cols to create dummy columns for race and age, then use polspline::polymars(y,x) to identify which parameters were most important. Confirmed that we should include admission stage, age, and race as predictors. It also showed us that we need to include age as a non-linear in our model - relationship changed as age increased linearly.
-   Included age as a polynomial to 2nd degree in regression formula. Tried out poly(age, 3) and poly(age, 4) too, but resulting p-values were less significant.
-   Try different transformations of outcome (LOS): Log, square root, cubic root.
-   Applied mgcv::gam with spline or smooth function (s(age)), which confirmed that we should use age as polynomial.
-   As of now, undecided on how to treat or parameterize race. Regression results show significance for Race-Other category, and we are not sure why this could be.
