---
title: "Clean UTSW Demographics and Covid Staging Data"
author: "Steph Reynolds (Stephanie.Reynolds@ucsf.edu)"
date: "`r format(Sys.time(), '%b %d, %Y  %H:%M %p')`"
output:
  html_document:
    toc: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Background

## Project 
MINDSCAPE: Modeling of infectious network dynamics for surveillance, control and prevention enhancement

## Description
This file cleans the demographics and daily Covid staging data (`dm_covid_stg_11.08.21.csv`) and prepares for analysis (e.g., create new variables and transform variables to proper data type). The resulting dataframe (`df`) is then de-duplicated to create another dataset where each row represents a unique patient (`one_row_per_pt`).

## Source Data
  - Demographics for Covid Patients (`COVID_Reg_Hosp_Demographics_03012020_123121_FINAL.csv`)
  - Ordinal Scale Data for Covid Patients (`UTSW_Ordinal_Scale_data_unt_12312021_FINAL.csv`)
  
# Load required packages 
```{r Load packages}
library(tidyverse)
library(forcats)
```

# Import and clean ordinal scale (staging) data
```{r Import and clean ordinal scale (staging) data}
stage_raw <- read_csv("/Users/sreynolds2/Documents/GitHub/MS-UTSW_Covid_Staging/data/raw/UTSW_Ordinal_Scale_data_unt_12312021_FINAL.csv")

stage <- stage_raw %>% 
    filter(!is.na(ordinal_scale_final)) %>% 
    select(PAT_ID, event_time, ordinal_scale_final, HSP_INITIAL_DATE) %>% 
    mutate(date = as.Date(event_time)) %>% 
    rename(ID = PAT_ID,
           stage = ordinal_scale_final) %>% 
    group_by(ID, date) %>% 
    mutate(stage = max(stage)) %>% 
    distinct(ID, date, stage)
```

# Import and clean demographic data 
```{r Import and clean demographic data }
demog_raw <- read_csv("/Users/sreynolds2/Documents/GitHub/MS-UTSW_Covid_Staging/data/raw/COVID_Reg_Hosp_Demographics_03012020_123121_FINAL.csv")

demog <- demog_raw %>%  
    distinct(PAT_ID, BIRTH_DATE, .keep_all = T) %>% 
    rename(ID = PAT_ID,
           BMI = Initial_BMI) %>% 
    mutate(age = as.numeric((as.Date(ADM_DATE) - as.Date(BIRTH_DATE)) / 365.25),
           sex = factor(Sex),
           race = factor(Race),
           ethnicity = factor(Ethnic_Group),
           zip = as.character(Zip_Code)) %>% 
    select(ID, age, sex, race, ethnicity, zip, BMI)
```

# Collapse levels for ethnicity and race variables
Make sure levels align with those in the UCSF dataset, so the two can be compared.
```{r Collapse levels for race and ethnicity vars }
# Collapse levels for ethnicity and race variables (aligning with UCSF dataset)
demog$ethnicity <- fct_collapse(demog$ethnicity,
                                "Not Hispanic or Latino" = c("Non-Hispanic/Latino", "White"),
                                "Unknown" = c("Unknown", "Declined"))

demog$race <- fct_collapse(demog$race,
                           "Other" = c("Some other race", "American Indian or Alaska Native"),
                           "Unknown" = c("Unavailable/Unknown", "Declined"))
```

# Merge daily staging and demographics tables, and rearrange order of variables
```{r Merge demog and staging datasets}
all_pt_days <- inner_join(stage, demog, by = "ID") %>% 
    relocate(date, .before=age) %>% 
    relocate(stage, .before=age)
```

# Group by ID then create new variables and assign to `all_pt_days` 

## Variables include:

- LOS
- max_stage
- date_adm
- date_disc
- stage_adm
- stage_disc
- end_in_death
```{r Mutate variables: LOS, max_stage, date_adm, date_disc, stage_adm, stage_disc, end_in_death}
all_pt_days <- all_pt_days %>% 
    group_by(ID) %>% 
    mutate(stage = factor(stage,
                          ordered = TRUE,
                          levels = c("4", "5", "6", "7", "8", "9", "10")),
           LOS = as.integer(round(max(date) - min(date), 1)),
           max_stage = factor(max(stage), 
                              ordered = TRUE, 
                              levels = c("4", "5", "6", "7", "8", "9", "10")), 
           date_adm = min(date),
           date_disc = max(date),
           stage_adm = factor(stage[date==min(date)],  # could also use as.integer to transform stage variables to integer type 
                              ordered = TRUE,
                              levels = c("4", "5", "6", "7", "8", "9")),
           stage_disc = factor(stage[date==max(date)], 
                               ordered = TRUE, 
                               levels = c("4", "5", "6", "7", "8", "9", "10")),
           end_in_death = factor(ifelse(max_stage==10, 1, 0)))

# Save resulting df as CSV and RDS
write_csv(all_pt_days, "/Users/sreynolds2/Documents/GitHub/MS-UTSW_Covid_Staging/data/dm_stg_all_pt_days.csv")
saveRDS(all_pt_days, file="/Users/sreynolds2/Documents/GitHub/MS-UTSW_Covid_Staging/data/dm_stg_all_pt_days.rds")

```

# Remove duplicates to create df that contains one row per unique patient
```{r}
unique_pts <- all_pt_days %>% 
    distinct(ID, .keep_all = T) %>% 
    select(-c(date, stage))

# Save resulting df as CSV and RDS
write_csv(unique_pts, "/Users/sreynolds2/Documents/GitHub/MS-UTSW_Covid_Staging/data/dm_stg_unique_pts.csv")
saveRDS(unique_pts, file="/Users/sreynolds2/Documents/GitHub/MS-UTSW_Covid_Staging/data/dm_stg_unique_pts.rds")

```

